---
title: "MEDEG 2021-2022: Thesis Working Draft"
subtitle: 'Co-ethnic concentration, social capital, and economic integration of immigrants in Spain (1998-2007)'
author: "Madeline Swarr"
date: "07 July 2022"

header-includes:
    - \usepackage{setspace}\singlespacing
    - \usepackage[bottom]{footmisc}
    - \usepackage{float}
geometry: "left=1.7cm,right=1.7cm,top=1.8cm,bottom=1.8cm"
output: html_document
bibliography: references.bib
csl: apa-5th-edition.csl
nocite: "@*"
fontsize: 12pt
---

**Last updated: `r format(Sys.time(), '%d %B %Y')`**

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
######################## Specify head directory ################
head_dir <- "/Users/papapeach/Documents/UC3M/Thesis/Data Analysis/ENI-Analysis"
run_date <- "06jul2022"

######################## (Down)load libraries ######################
dependency_dir <- paste0(head_dir, "/dependencies")
source(paste0(dependency_dir, "/get_packages.R"))
get_packages(c("plyr", "tidyverse", "ggplot2", "RColorBrewer", "openxlsx", "sf", "terra", "spData", 
               "spDataLarge", "mapSpain", "tmap", "reshape", "gifski", "RColorBrewer", "kableExtra",
               "MASS", "foreign", "Hmisc", "reshape2", "dplyr", "stargazer", "robustbase"))
```

## Abstract
Social science has always been concerned with understanding the meso-level forces that connect individuals to the whole. In the case of the ethnic immigrant, this integration into a new society, characterized by differences in institutions, language, culture, and socioeconomic structure, is a long and arduous process that those of the native population do not always face, and if they do, do not face as intensely. 

This paper defends the idea that the concentration of immigrants in co-ethnic geographical clusters, while intuitively helpful in protecting against the psychological shocks of such displacement, can also help explain why those living in certain "ethnic enclaves", as this clustering phenomenon is often referred to, experience different economic outcomes depending on the specific enclave one belongs to. This analysis uses the 2007 Spanish National Survey of Immigrants and Spanish registry data to empirically assess the relationship between residential co-ethnic density and economic well-being and job precarity of immigrants who arrived to Spain between the years 1998 and 2007. 

## Introduction
In immigration studies, the concept of “ethnic penalty” illustrates the fact that a meritocratic society does not exist for minorities, and that the degree of penalty on labor mobility and earnings depends on the ethnic group being referenced [@hasmath2012ethnic]. Therefore, the classic determinants of economic well-being for ethnic immigrants do not have the same effects as for those of the native population nor as for non-ethnic immigrants. As Kuznets puts it, “all minorities are characterized, at a given time, by an occupational structure distinctly narrower than that of the total population and the majority” [@kuznets1960economic]. 

This occupational structure is most rigid for immigrants in the years directly following arrival in the destination country, when little opportunity has been given to accumulate and assimilate the human capital necessary to participate fully in the local labor market and legal and language barriers are aplenty [@chiswick1978]. Given the societal and legal restrictions placed on immigrants, the impact that social networks have on their socioeconomic integration becomes that much greater. Even if social ties are not leveraged directly for economic gain, the influence that pertaining to an "in-group" has on protection against discrimination, information sharing, and trust can indirectly affect one's skill acquisition, job prospects, salary, and general well-being. This influence is evidenced by the long history of geographical concentration of immigrant groups by ethnicity [@chiswick2001determinants] [@kerr2015social]. 

While much of the debate surrounding economic assimilation has centered around measuring the rate of convergence between native and immigrant wages over time, more recent research has focused on evaluating the determinants of differences in economic assimilation within cohorts of immigrants, specifically the effects of co-ethnic clustering depending on national origin, gender, sector of work, and legal status, and whether these effects are positive or negative depending. This paper will attempt to shed further light on these potential effects on salaries and job precarity of recent immigrants to Spain during the 10-year period between 1998 and 2007, utilizing detailed micro data from 15,465 individuals surveyed in the 2007 National Survey of Immigrants (*Encuesta nacional de inmigrantes*, hereon out referred to as the ENI) and linking them to population data from the municipal registry database (*Estadística del Padrón continuo*) to draw correlations between these economic outcomes and localized co-ethnic concentration, while controlling for demographic and human capital determinants.

[Insert main findings here.]

### Theoretical Framework
As developed in the late 1960s and early 1970s by a group of young Marxist economists, dual labor market theory states that developed economies are split into two sectors – the primary sector characterized by unions, good working conditions, and job security, and the secondary sector characterized by temporary contracts, minimum wage, poor working conditions, and little opportunity for promotion (see @doeringer1975unemployment for more). As the demographic and sectoral transition continues in the receiving country, a growing number of skilled jobs will go disproportionately to the native and educated population while vacated unskilled jobs go to those lacking competitive skills, often times those being minorities and immigrants, creating a labor market imbalance and segmentation that is hard to break out of [@harrison1979theory]. 

The importance of social capital among immigrants for correcting for information imbalances, resource scarcity, cultural displacement, restricted rights, and discrimination has been documented extensively in the literature (see @massey1987social, @munshi2003networks, and @epstein2010migration for examples). Many fertile lines of research have shown that geographic and sectoral clustering of immigrant groups by ethnicity can provide economic support when it is lacking elsewhere, as well as breed productivity and innovation through information-sharing and skills acquisition (see @light1990korean, @ozgen2012immigration, and @kerr2015social for examples). On the other hand, it is also argued that ethnic enclaves can restrict job mobility and reduce incentives to build skills (like language ability) depending on how segregated the enclave is from the surrounding labor market [@borjas1994ethnicity].

The debate continues about whether the role of the ethnic enclave is beneficial to the economic well-being of an individual, and if so, for where, for what, and for whom in particular. @portes1985latin have studied extensively the situation of Cuban immigrants in Miami, Florida and have generally found greater returns for Cuban-owned businesses and their co-ethnic employees than for their counterparts in secondary and primary labor markets, while @sanders1987limits have argued that it is only the business owners who benefit from the solidarity and insularity of the enclave. @barry2005enclaves show that ethnic clustering in the U.S. is associated with lower earnings amongst male immigrants from non-English speaking countries, along with lower English attainment. @kanas2012social provide evidence that that residential co-ethnic concentration in Germany has a weak negative effect on the economic returns of education achieved in the host country.

There are many hypotheses about the mechanisms that may explain the differences in additive effects of ethnic concentration on such economic outcomes. For example, distinguishing economic returns to human capital between that education achieved in the origin country and that education achieved in the host country, we may expect greater returns to the former in an ethnic enclave where that human capital is valued more. When looking at the interaction between ethnic concentration and gender, given the greater constraints on labor market integration of female immigrants in comparison to male immigrants, we may expect fewer returns to the ethnic enclave due to less opportunities to leverage that social capital for the improvement of one's economic or laboral situation. There may be certain sectors or businesses that benefit from trust, cooperation, and shared norms or cultural values more so than others, also potentially leading to differences in economic returns to ethnic concentration for, say, a vendor selling ghee versus a vendor selling hamburgers in a predominantly Hindu neighborhood.

### Contextual Motivation
While many studies on ethnic enclaves and economic well-being have been situated in the United States, United Kingdom, and Germany, literature looking at the Spanish context has been rather sparse despite the fact that Spain is home to one of the largest immigrant populations in the world.^[As measured by percentage of the resident population.] According to the IOM’s 2020 World Migration Report, Spain was Europe's sixth most popular migrant destination in 2019 [@wmr2020]. Spain as a major destination for international migration is a trend that began around the turn of the twentieth century, with now at least 15% of the national population consisting of foreign nationals, with Morocco, Romania, United Kingdom, and Colombia representing the top countries of origin as of the second quarter of 2021 [@inepressrelease].^[These estimates vary due to the counting issues that arise from unregistered immigrants, as well as how an immigrant is defined. For example, an immigrant may be defined as anyone who is foreign-born (who may or may not have Spanish nationality), or anyone who possesses foreign citizenship.]

```{r echo = FALSE, warning = FALSE, message = FALSE, animation.hook="gifski", results='hide', fig.keep='all', fig.align = "center"}
####################    Padrón directory specification   #######################
padron_dir <- paste0(head_dir, "/Data/" )
padron_data_name <- "población_provincia_serie_temporal.xlsx"
output_dir <- paste0(head_dir, "/output/", run_date, "/4_padron_prepared")
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = T)
}

####################   Prepare and merge geo data #########################
## Population density attributes by province and country of origin
dat_padron <- read.xlsx(paste0(padron_dir, "/", padron_data_name), sheet = 3, startRow = 8, cols = c(3:28)) 
## Filter to all immigrants, reorder columns, and convert wide to long to be able to facet by year
dat_padron_all_melt <- dat_padron %>% filter(Origin == "TOTAL EXTRANJEROS" & Destination_mapSpain != "Total") %>%
  dplyr::select(c(Destination_mapSpain, Origin, `1998`:`2021`)) %>% melt(., id.vars = c("Destination_mapSpain", "Origin")) %>%
  dplyr::rename(density = value, year = variable)
## Get province shapes
provinces <- esp_get_prov_siane()
provinces_names <- levels(factor(provinces$ine.prov.name))
## Merge attributes with geo data
dat_geo <- sp::merge(provinces,
                    dat_padron_all_melt,
                    by.x = "ine.prov.name",
                    by.y = "Destination_mapSpain",
                    all.x = TRUE
)

################### Plot with tmap #####################
## Plot immigration population density by province over years 1998 - 2021
ethnic_density_map_anim <- tm_shape(dat_geo) +
  tm_polygons("density", palette = "Reds", style="cont", n = 10, 
              title = "Immigrant density \n (% of population)",
              id = "ine.prov.name") +
  tm_facets(along = "year", free.coords = FALSE) +
  tm_legend(position = c("left", "top")) +
  tm_credits(text = "Data source: INE's padrón continuo", position = c("right", "bottom"))
ethnic_density_map_anim

```
\vspace{3truemm}

There is a plethora of research documenting the changes of internal migration rates within Spain as well, with overwhelming evidence showing that foreigners are much more likely to move around the country than natives are in response to labor market demands, and that they are indeed meeting those demands with 45% of new employment being filled by immigrants in 2019 [@domenech2020]. This is a rather intuitive yet important finding considering that having high labor mobility is essential in fostering a robust and dynamic economic system able to respond quickly to shocks in supply and demand [@blanchardkatz1992]. The rather high prevalence of temporary contracts in Spain paired with concerns of growing systematic unemployment, stagnation in innovation, regional disparities in unemployment rates, and an aging population make it critical that immigration not be viewed as a burden, but rather as an opportunity to integrate young, diverse and dynamic labor into the Spanish workforce [@liu2018regional] [@roses2020accounting]. 

## Data  
### *Encuesta nacional de inmigrantes* (ENI) from 2007
Principal data for this analysis consists of micro-level data from the Spanish National Institute of Statistics' (INE) National Survey of Immigrants (ENI). The survey consists of 15,465 immigrant households living in Spain as of 2007 that were randomly selected from predesignated pools of households with at least one individual registered as living there who met the following requirements: The individual 1) was born outside of Spain; 2) was 16 years or older at time of surveying; 3) had established residency in Spain for at least one year prior to surveying and if not the case, had the intention of staying in Spain for at least a year; and 4) for those who had Spanish nationality from birth (but born outside of Spain), must had been at least 2-years-old at the time of arrival to Spain. 

The sampling process was defined at the autonomous community level on differing population strata to target certain groups of nationalities of interest, particularly those from Ecuador, Morocco, Romania, Latin American countries excluding Ecuador, African countries excluding South Africa, Asian countries excluding Japan, North American excluding Mexico and Oceania, and the EU15 and European Economic Area excluding EU member countries. 87.37% of the desired sample size was achieved (15,465 of 17,700), with 57.62% of surveyed households coming from the primary targeted population and 42.38% coming from the supplementary targeted population. Household surveys were carried out between November of 2006 and February of 2007.

According to the INE's report on the ENI (or NIS as referred as below), 

> "The National Immigrant Survey conducted in 2007 is aimed at studying the demographic and social characteristics of those persons born abroad, as well as their migratory itineraries, work and residential history, family relations and relations with their country of origin and with the Spanish society. NIS-2007 places a special emphasis on the study of social relations and personal networks linked to the act of migrating. The survey has a general purpose and a very broad subject matter, and is the first large survey on the recent migratory flows, covering the entire national territory and all immigrant groups." [@reher2008informe].^[Information on municipality of residence in Spain and history of movements between municipalities is available, but fictitious names of the municipalities are assigned for data protection reasons in accordance with the *Secreto Estadístico* of Spain. I am currently in the process of requesting de-anonomized data to be able to map ENI data onto municipal-level registry data.]

**Link to the raw data, dictionaries, survey methodology, and questionnaire used to collect responses can be found [here](https://www.ine.es/dyngs/INEbase/es/operacion.htm?c=Estadistica_C&cid=1254736177005&menu=resultados&idp=1254735573002#!tabs-1254736195389 "ENI").** 


```{r echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
## Set directory where processed ENI data can be found
input_dir <- paste0(head_dir, "/Output/", run_date, "/1_processed")
## Load processed ENI data
this_data <- readRDS(paste0(input_dir, "/datos_eni_07_processed_subset.rds"))
this_data_all <- readRDS(paste0(input_dir, "/datos_eni_07_processed.rds"))

## Restrict to immigrants who immigrated for economic reasons and came in 1998 or later
this_data_restricted <- this_data_all %>% filter(economic == 1 & arrival_spain >= 1998)
## Which countries are most represented?
countries <- table(this_data_restricted$country) %>% as.data.frame(responseName = "Freq") %>% 
  dplyr::rename(country = Var1) %>% arrange(desc(Freq))
these_top_countries <- countries %>% filter(Freq > 175)

## Year of arrival by most represented countries of origin
this_data %>% filter(country %in% these_top_countries$country) %>% ggplot(aes(x=arrival_spain, fill=country)) +
  geom_histogram(bins = 60) +
  scale_fill_brewer(palette="Set2") +
  labs(title = "Counts of immigrants in ENI by arrival year (1945 - 2007) \n Top represented countries (economic immigrants)",
       x = "Arrival Year",
       y = "Count",
       fill = "Country of origin") +
  scale_x_continuous(limits = c(1945, 2007)) +
  theme_bw()

```

### *Estadística del Padrón continuo*
Aggregate population data by year and Spanish province will be analyzed from the municipal registry database (*Estadística del Padrón continuo*) in order to map province-level yearly population data to immigrants in the ENI by country of origin. Registry data is available from 1998 onwards at the province-level (52 provinces), and from 2003 onwards at the municipal-level (8,112 municipalities). Spanish provinces are measured at the NUTS-3 level. Country of origin is measured at the GADM-1 level. Registry data for certain countries of origin with low representativeness is not available at an individual country-level, and thus must be grouped by regional categories.^[I am still in the process of consolidating these regional categories.] Municipal-level registry data is also available from 1997-2002, but it is stratified at the continent of origin level.

**Link to the aggregated data files with registry information counted by year, province, and nationality can be found [here](https://www.ine.es/jaxi/Tabla.htm?path=/t20/e245/p04/provi/l0/&file=0ccaa002.px&L=0 "Padrón continuo").** 

```{r echo = FALSE, warning = FALSE, message = FALSE, results='hide', fig.keep='all', fig.align = "center", fig.height = 8, fig.width = 10, strip.white=T}
## Filter population data by province to the top represented countries of origin and to year 2006
dat_padron_top_2006 <- dat_padron %>% 
  dplyr::filter((Origin %in% these_top_countries$country) & (Destination_mapSpain != "Total")) %>%
  dplyr::select(c(Destination_mapSpain, Origin, `2006`))
dat_padron_top_2006_melt <- melt(dat_padron_top_2006, id.vars = c("Destination_mapSpain", "Origin")) %>%
  dplyr::rename(density = value, year = variable)
## Get province shapes
provinces <- esp_get_prov_siane()
provinces_names <- levels(factor(provinces$ine.prov.name))
## Merge attributes with geo data
dat_geo_top_2006 <- sp::merge(provinces,
                    dat_padron_top_2006_melt,
                    by.x = "ine.prov.name",
                    by.y = "Destination_mapSpain",
                    all.x = TRUE
)

################### Plot with tmap #####################
## Population density in 2006 for each country of top origin
ethnic_density_top_2006 <- tm_shape(dat_geo_top_2006) +
  tm_polygons("density", palette = "Reds", style="cont", n = 10, 
              title = "Immigrant density \n (% of population)",
              id = "ine.prov.name") +
  tm_facets(by = "Origin", free.coords = FALSE, nrow = 4, ncol = 2) +
  tm_layout(asp=0)
ethnic_density_top_2006
```

## Analysis

### Study Population
To account for potential endogeneity stemming from systematic differences in economic motivation amongst immigrants, the population will be restricted to those deemed as “economic immigrants”, or those who have come to Spain “*por falta de empleo*” ($FEMPXX$), “*para buscar un empleo mejor*” ($BEMPXX$), or “*por cambio de destino laboral*” ($DSLAXX$).^[*Tabla 2.1* of ENI questionnaire documents individual’s migration histories, with the last entry representing his/her decision-making when immigrating to Spain.]

On top of the population of economic immigrants, we will restrict to those who arrived in 1998 or later since registry data at the province-level by country of origin are only available from 1998 onwards ($ALLEXX$). 

Finally, to maximize granularity of population data and minimize potential bias due to recall error, robustness checks will be implemented by restricting the population to those who had arrived to Spain in 2003 or later, since registry data at the municipal-level by country of origin are available from 2003 onwards, and a maximum recall window of four years is a reasonable window to minimize the effects of memory decay, while also maintaining a robust sample size.

```{r echo = FALSE, warning = FALSE, message = FALSE}
##################### POPULATION RESTRICTION ATTRITION #################
restriction_data <- this_data %>% dplyr::select(economic, arrival_spain) %>% 
  mutate(economic = ifelse(economic == 1, 1, 0), 
         economic_1998 = ifelse(economic == 1 & arrival_spain >= 1998, 1, 0),
         economic_2003 = ifelse(economic == 1 & arrival_spain >= 2003, 1, 0)) %>% 
  dplyr::select(-arrival_spain)

restriction_stats <- data.frame(yes = colSums(restriction_data ==1, na.rm = T),
                                no = colSums(restriction_data==0, na.rm = T)) %>%
mutate(total = no + yes,
       percent = (yes/total)*100)


rownames(restriction_stats) <- c("Economic?", 
                                 "Economic and arrived post-1998?",
                                 "Economic and arrived post-2003?")

## Output restriction stats
kable(restriction_stats, format = "html", digits = c(0, 0, 0, 2), row.names = T, align = 'c', 
      col.names = c("Yes", "No", "Total", "% of Total (Yes)"),
      caption = "ENI Population Restrictions") %>%
  kable_styling()

```

### Modelling
To test whether social capital, $SC$, has an effect on economic outcomes, $Y$, of an ethnic immigrant $i$ from country $c$ living in region $r$ at time of surveying, I will utilize the following theoretical model:

$$Y_{i,c,r} = \beta_0 + \beta_1SC_{i,c,r} + \bf{\beta_x}x_{i,c,r} + \epsilon $$
According to the work of @carliner1980wages that compares the wages and hours worked of immigrants to the U.S. across three generations, in addition to the classic Mincer determinants such as education ($EDUC$) and prior work experience ($EXP$), human capital can be proxied for by time spent in the destination country ($YEARS$) as one goes accustoming to the new culture, norms, and general way of life in which he finds himself. Controls will also be included for age ($AGE$), the square of age ($AGE^2$), and sex ($SEX$).^[A control for whether one has a spouse or family will also be considered.] Additionally, location-based fixed effects such as language, culture, and other time-invariant omitted variables are proxied for using home and destination locations ($COUNTRY$, $REGION$).^[Due to high multicollinearity between $REGION$, $COUNTRY$, and the measure of co-ethnic density, more aggregate or variable measures of $COUNTRY$ and $REGION$ will be considered, like urban vs. rural or the depravity index to measure $REGION$, and continent of origin instead of $COUNTRY$.] Time-variant controls will be investigated to try to account for changing job markets over time by location when looking at response variables measured in different years (time it took to find a first job, for example, which depends on the year in which the immigrant arrived).^[The deprivation index for small areas in Spain may be controlled for as has been done in @becares2015social [@sanchez2008deprivation].] Finally, one’s legal status ($PERMIT$) plays a significant role in determining economic and social outcomes, as well as whether one was even observed in the ENI data or answered truthfully or accurately the questions of the survey. Having work/residency authorization prior to arrival in Spain is an indicator of those who came with pre-contracted work and thus may indicate systematic differences arising from selection bias of this population.^[In 1985, the General Regime was established to grant employers visa quotas that they could use to sponsor foreign workers. Broad reforms of the degree took place during the 2000s that were effective in creating a demand-driven system focused on filling the fastest growing industries [@imopolicies].] 

#### Response Variables
**Earnings and wages:** The principal response variables to be measured will deal with earnings, specifically monthly earnings and hourly earnings. Monthly earnings variable can either be answered with a real value ($INGRE$) or by a given range ($INGRTR$), but an analysis using salary ranges will only be conducted in the case of a low-powered sample. Hourly earnings ($EARN$) will be calculated by dividing monthly earnings ($INGRE$) by hours worked a week ($HORAS2$) multiplied by four to get hours worked habitually a month. When measuring earnings and wages, the population will be restricted to those individuals who work at least 30 hours a week or more habitually.

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.height = 7.0, fig.width = 9}
this_pop <- this_data_restricted %>% filter(country %in% these_top_countries$country)
## Hourly earnings
mu_hour_earn <- ddply(this_pop, c("sector2", "country"), summarise, grp.mean=mean(hour_earn, na.rm = T)) 
mu_hour_earn_all <- ddply(this_pop, c("country"), summarise, grp.mean=mean(hour_earn, na.rm = T)) %>%
  mutate(sector2 = "All")

## Hourly earnings distribution by sector of work, facetted on country of origin
gg_hour_earn <- this_pop %>% mutate(sector2_trim = gsub(";.*", "", sector2)) %>%
  ggplot(aes(hour_earn, fill = sector2_trim, color = sector2_trim)) + 
  geom_histogram(stat = "bin", bins = 40) +
  labs(title = "Counts of immigrants by hourly earnings reported and sector of activity",
       x = "Estimated hourly earnings",
       y = "Count",
       fill = "Sector of activity") +
  geom_vline(data=mu_hour_earn, aes(xintercept=grp.mean, color=sector2),
             linetype="dashed") +
  geom_vline(data=mu_hour_earn_all, aes(xintercept=grp.mean), color = "black", linetype = "solid") +
  theme(legend.position="right") + 
  facet_grid(country~., scales = "free") +
  scale_x_continuous(limits = c(0, 18)) +
  guides(color = F)
gg_hour_earn
```

**Job precarity:** Secondary response variables of interest will look at employment and unemployment situations, specifically the time it took to find a first job upon arriving to Spain ($TPO1$) and the number of unemployment periods experienced for a month or more ($NVPARO$), both measured on an ordinal scale.

#### Explanatory Variables
Recent research has stressed the characterization of social capital to better identify the mechanisms through which it can affect the economic well-being of an individual. While there are many interpretations of what constitutes social capital, I will extend my research to include proxies for the following three principal characteristics that have been discussed in the literature – trust and cooperative norms, personal relationships, and civic engagement – without any claims that these are indeed the causal mechanisms behind observed differences in wages and job precarity (see @scrivens2013four for an overview).

Trust and cooperative norms will be proxied for by the concentration of co-ethnics living in a given province/municipality, defined as the proportion of the population of the same country of origin that were registered as of January 1, 2006 as living in the same Spanish province/municipality as the immigrant of reference. To address the potentially endogenous decision to move to a certain province/municipality where one *expects* to be able to leverage co-ethnic networks for economic well-being, I will assess the possibility of instrumenting "concentration of co-ethnics" for country of origin *c* in region *r* in 2006 with corresponding data from the earliest year available as well as assess the probability of changing residence due to regional differences in employment or wage level.

Following @bolibar2020social, the relationship between immigrant social networks and economic well-being will be investigated to see if there is any effect of personal relationships, and to what extent the characterization of these ties (co-ethnic familial ties versus ties with Spanish natives) and the geographical distance of each may explain differences in the effect. 

Following @tolbert1998local and theories of civil society, I will additionally use information on civic engagement through participation in local NGOs, religious groups, extracurricular classes, or municipal elections to assess any association with positive economic outcomes. There are endogeneity concerns that will be addressed later when studying these potential explanatory variables and their effect on wages, considering the probability of civic participation is highly correlated with socioeconomic and educational background.  

```{r echo = FALSE, warning = FALSE, message = FALSE}
## Filter to the principal study population (economic immigrants that arrived in 1998 or later)
study_pop <- this_data_all %>% filter(economic == 1 & arrival_spain >= 1998)

################### Civic engagement and participation ######################
participation_data <- study_pop %>%
  dplyr::select(starts_with(c("PNX", "PX", "PELEC"))) %>% filter(complete.cases(.))

## Select anyone that responded to civic engagement and participation questions
participation_data_all <- study_pop %>% 
  dplyr::filter(!is.na(participation_index)) %>%
  dplyr::select(participation_indicator)

## Stats by variable
civic_engagement_stats <- data.frame(yes = colSums(participation_data==1, na.rm = T),
                                     no = colSums(participation_data==0, na.rm = T)) %>%
  mutate(total = nrow(participation_data),
         percent = (yes/total)*100)

## Stats by any variable                                  
civic_engagement_stats_any <- data.frame(yes = sum(participation_data_all==1),
                                         no = sum(participation_data_all==0)) %>%
  mutate(total = nrow(participation_data_all),
         percent = (yes/total)*100)

## Append stats by variable and stats by any variable
civic_engagement_stats_all <- rbind(civic_engagement_stats, civic_engagement_stats_any)
rownames(civic_engagement_stats_all) <- c(rownames(civic_engagement_stats), "Any participation")

## Output stats
kable(civic_engagement_stats_all, format = "html", digits = c(0, 0, 0, 2), row.names = T, align = 'c', 
      col.names = c("Yes", "No", "Total", "% of Total (Yes)"),
      caption = "Civic engagement variables and participation counts by ENI variables") %>%
  kable_styling() %>%
  footnote(general = "**PX---**,**PNX---** variables answer the question, *¿Participa en algunos de los siguientes grupos o asociaciones orientados específicamente o no a extranjeros?*, while **PELEC** answers, *¿Ha participado en alguna elección municipal en España?*")
```

### Hypothesis Testing
Using the standard Mincer equation as a basis, the multiple linear regression models estimating the log of monthly earnings and the log of hourly earnings ($logEARN$) for an individual ethnic immigrant to Spain at time of surveying will be the following:

\begin{equation*}
  \begin{aligned}logEARN_{i,c,r} = \beta_0 + \beta_1SC_{i,c,r} + \beta_2EDUC_i + \beta_3EXP_i + \beta_4LANG_i + \beta_5YEARS_i + \beta_6AGE_i + \beta_7AGE^2_i + \beta_8SEX_i + \beta_{9}COUNTRY_c + \\ \beta_{10}REGION_r + \beta_{11}PERMIT_i + \epsilon_{i,c,r}
    \end{aligned}
\end{equation*}

When $SC$ represents co-ethnic density that corresponds to immigrant $i$ from country $c$ living in region $r$, the coefficient of interest, $\beta_{1}$, represents the difference in log of monthly earnings or log of hourly earnings for someone with one percentage point more of the population in the province in which he is living made up of co-ethnics, *ceteris paribus*.^["Co-ethnic" is defined as someone with the same nationality as the reference individual.] 

In the simplest case for which $SC$ is a binary indicator of whether someone possesses at least one of the indicators for social capital, the coefficient of interest, $\beta_{1}$, represents the baseline difference in log of monthly earnings and log of hourly earnings between someone who possess social capital ($SC = 1$) versus a similar someone who does not ($SC = 0)$.

It is plausible that amongst immigrants of the same ethnic group there will not be much variation in the sector in which they work. To understand the effect of social capital given that sectoral choice is often limited for ethnic immigrants, it might be more appropriate to examine the interaction between social capital and job sector and its effect on wages. Therefore, the linear models we would be more interested in estimating would be the following: 

\begin{equation*}
  \begin{aligned}
logEARN_{i,c,r} = \beta_0 + \beta_1SC_{i,c,r} + \beta_2SECTOR + \beta_3SC \times SECTOR + \beta_4EDUC_i + \beta_5EXP_i + \beta_6LANG_i + \beta_7YEARS_i + \\ \beta_8AGE_i + \beta_9AGE^2_i + \beta_{10}SEX_i + \beta_{11}COUNTRY_c + \beta_{12}REGION_r + \beta_{13}PERMIT_i + \epsilon_{i,c,r}
  \end{aligned}
\end{equation*}

The hypothesis would then be that social capital carries differing effects for certain occupational sectors (and thus for certain immigrant groups), possibly carrying more effect for those sectors where information-sharing, trust, and collaboration are more important for productivity. Once again taking the naïve social capital dummy variable, we would test this hypothesis by looking at the effect, $\beta_3$, of the interaction between social capital and occupation sector of employment $SC \times SECTOR$. If there are three sectors, for example, then the term $\beta_3 SC \times SECTOR$ would be expanded as such:

$$\beta_3SC \times SECTOR = \beta_{3,1}SC \times SECTOR_1 + \beta_{3,2}SC \times SECTOR_2 + \beta_{3,3}SC \times SECTOR_3$$
with $SECTOR_i$ as a binary indicator of whether the person works in $SECTOR_i$ or not. We would test the following null hypothesis to see whether there is an additive effect of social capital and occupational sector:

$$H_0: \beta_{3,1} = \beta_{3,2} = \beta_{3,3} = 0$$
and the following null hypothesis to see whether the additive effects of social capital and occupational sector differ by occupational sector:

$$H_0: \beta_{3,1} \ne \beta_{3,2} \ne \beta_{3,3} = 0$$
Other interactions will be tested as well, specifically ones where the effect of having more social capital are expected to be more advantageous (or disadvantageous) for certain groups over others. These interactions could be by gender ($SEX$), by country of origin ($COUNTRY$), or by work authorization status ($PERMIT$).

### Preliminary Regressions
```{r echo = FALSE, warning = FALSE, message = FALSE}
# Restrict to primary study population 
this_pop <- this_data_restricted %>% dplyr::filter(!is.na(log_hour_earn)) %>% mutate(eth_dens_prov_sq = eth_dens_prov^2)
this_pop$spanish <- factor(this_pop$spanish, levels = c("Necesita mejorar", "Suficiente", "Bien", "Nativo"))
this_pop$sex <- factor(this_pop$sex, levels = c("Male", "Female"))

#### Models for Table 1 #####
ethn1 <- lm(log_hour_earn ~ eth_dens_prov, data = this_pop)
supp1 <- lm(log_hour_earn ~ any_support, data = this_pop)
part1 <- lm(log_hour_earn ~ participation_indicator, data = this_pop)
# With covariates
model5 <- lm(log_hour_earn ~ eth_dens_prov + any_support + participation_indicator + MNIV + spanish + years + age + age_sq + sex + work_permit, data = this_pop)
model6 <- lm(log_hour_earn ~ eth_dens_prov +  any_support + participation_indicator + MNIV + spanish + years + age + age_sq + sex + work_permit + sector2, data = this_pop)
model7 <- lm(log_hour_earn ~eth_dens_prov +  any_support + participation_indicator + MNIV + spanish + years + age + age_sq + sex + work_permit + sector2 + occupation2, data = this_pop)
model8 <- lm(log_hour_earn ~ eth_dens_prov +  any_support + participation_indicator + MNIV + spanish + years + age + age_sq + sex + work_permit + sector2 + occupation2 + country, data = this_pop)
model9 <- lm(log_hour_earn ~ eth_dens_prov +  any_support + participation_indicator + MNIV + spanish + years + age + age_sq + sex + work_permit + sector2 + occupation2 + country + municipality, data = this_pop)

```


```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', resize.height=4,resize.width=11}
#### Table 1 ####
stargazer(model5, model6, model7, model8, model9,
          type="html",
          title = "Effect of social capital variables on log of estimated hourly earnings, with covariates",
          #colnames = FALSE,
          #column.labels = c("Total"),
          #column.separate = c(3, 5),
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Co-ethnic concentration", "Support upon arrival", "Civic engagement", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Years in Spain", "Age", "Age squared", "Female", "Work authorization?"),
          keep = ,
          omit = c("sector2", "occupation2", "country", "municipality"),
          add.lines=list(c('Sector controls', "No", "Yes", "Yes", "Yes", "Yes"),
             c('Occupation controls', "No", "No", "Yes", "Yes", "Yes"),
             c('Country of origin controls', "No", "No", "No", "Yes", "Yes"),
             c('Municipality controls', "No", "No", "No", "No", "Yes")),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = F,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "footnotesize",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")

```
#### Observations
1. There is a significant negative effect of co-ethnic concentration in models (1) through (3) at $\alpha \le 0.5, 0.5, 0.1$, respectively). The model does not provide better fit when the effect is modeled as quadratic (not shown).

2. The negative effect of having friends or family to direct to upon arriving in Spain on the log of estimated hourly wage seen in models (1) through (3) gets absorbed with the addition of country of origin controls. We see a drastic change in effect size and sign when controlling for municipality of residence in model (5) .

3. In models (1) through (4), we see a positive and significant effect of civic engagement on wages. The high magnitude effect (on the order of a ~6.1% average increase in wages for those who have participated in a local organization or voted in a municipal election versus those who haven't) decreases as we control for more covariates. When controls for education and years since arrival are added (not shown), we see that the ones who are highly educated and have been in Spain more years are more likely to have participated in a local organization or voted in municipal elections, hence the decrease in the positive effect of civic engagement. 

4. Level of Spanish, which is hard to measure and a strong predictor of country of origin, might be better captured by years lived in Spain and country of origin.

5. Work authorization is expectedly a strong, positive predictor of hourly wages.

#### Issues  
1. Generally there is a lot of collinearity between the variables and not much sample size. 

2. There's not much variation in country of origin for high co-ethnic concentration levels (multicollinearity). Alternatively put, the most highly "ethnic" provinces are represented mainly by a few of the largest sending countries (Romania, Morocco, Ecuador). Thus, we don't know if we are controlling for country of origin or co-ethnic density, which are measured at different levels. 

3. Non-randomness of immigration and settlement patterns. Endogeneity from the decision to live in a certain zone, which is likely to be decided upon with economic-related considerations in mind. 

#### Next steps
1. Find more variation in the ethnic concentration measure by matching population data by year, country of origin, province of residence, and gender.

2. Perhaps consider the restriction of having to have immigrated for economic reasons to be used as a robustness check to enhance sample size. 

3. Drop Spanish ability as a predictor. 

4. Interaction terms (civic engagement and gender, ethnic density and sector or country of origin).

5. Instrument co-ethnic density with earliest ethnic density data available.

6. Look into other response variables – social networks and unemployment and labor mobility? For those who have a lot of ethnic social support upon arriving (but no pre-arranged job), how likely were they to find a first job quickly or
be in PARO? For those who got their first job through contacts, how likely were they to upgrade their jobs


\newpage

## References
<div id="refs"></div>

\newpage

## Code for this report  
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```




