---
title: "ENI Analysis: Preliminary regressions"
author: "Madeline Swarr"
date: "4/28/2022"
output: pdf_document
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[bottom]{footmisc}
    - \usepackage{float}
classoption: landscape
---

```{r echo = FALSE, warning = FALSE, message = FALSE}
########### ESPECIFICAMOS EL DIRECTORIO DEL CODIGO ################
head.dir <- "/Volumes/GoogleDrive/My Drive/Thesis/Data Analysis/ENI_ANALYSIS"
run_date <- "26abril2022"

# Descargar libraries 
dependency.dir <- paste0(head.dir, "/Pipeline/dependencies")
source(paste0(dependency.dir, "/get_packages.R"))
get_packages(c("MASS", "foreign", "Hmisc", "reshape2", "dplyr", "stargazer"))

####################    Asignacion de parametros    #######################
# Recogemos la ruta del script que se esta ejecutando
input.dir <- paste0(head.dir, "/Output/", run_date, "/1_processed")
output.dir <- paste0(head.dir, "/Output/", run_date, "/3_modelled")
if (!dir.exists(output.dir)) {
  dir.create(output.dir)
}

this_data_subset <- readRDS(paste0(input.dir, "/datos_eni_07_processed_subset.rds"))
this_data_subset$spanish <- factor(this_data_subset$spanish, levels = c("Necesita mejorar", "Suficiente", "Bien", "Nativo"))
this_data_subset$sex <- factor(this_data_subset$sex, levels = c("Male", "Female"))

####################### Restrict to primary study population ##########################
this_pop <- this_data_subset %>% dplyr::filter(economic == 1 & arrival_year >= 1998)
```


```{r echo = FALSE, warning = FALSE, message = FALSE}
######### 1b. Effect of concentration of coethnics in same province on estimated hourly earnings
this_pop_1b <- this_pop %>% filter(!is.na(log_hour_earn) & !is.na(eth_dens_prov))
model_1bi <- lm(log_hour_earn ~ eth_dens_prov, data = this_pop_1b)
#+ country_short + years + age + age_sq + sex + education + spanish + comunidad + occupation1 + work_permit

model_1bii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq, data = this_pop_1b)
#+ country_short + years + age + age_sq + sex + education + spanish + comunidad + occupation1 + work_permit

model_1biii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish, data = this_pop_1b)

model_1biv <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit, data = this_pop_1b)

model_1bv <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit + age + age_sq + sex, data = this_pop_1b)

model_1bvi <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit + age + age_sq + sex + country_short, data = this_pop_1b)

model_1bvii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality, data = this_pop_1b)

model_1bviii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1, data = this_pop_1b)

model_1bix <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1 + occupation1, data = this_pop_1b)
```

```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', resize.height=4,resize.width=11}

stargazer(model_1bi, model_1bii, model_1biii, model_1biv, model_1bv, model_1bvi, model_1bvii,
          type="latex",
          title = "Effect of concentration of coethnics in same province on log of estimated hourly earnings",
          colnames = FALSE,
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Co-ethnic concentration", "Co-ethnic concentration squared", "Years since arrival", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Work authorization?", "Age", "Age squared", "Sex - female"),
          keep = ,
          omit = c("country_short", "municipality"),
          add.lines=list(c('Country of origin controls', rep("No", 5), "Yes", "Yes"),
                         c('Municipality controls', rep("No", 6), "Yes")),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = TRUE,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "footnotesize",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")
```

## Table 1

### Observations
1. There is a significant negative effect of co-ethnic concentration in models (1) through (6) at $\alpha \le 0.05$). The models provide better fit when the effect is modeled as quadratic, and the coefficients indicate a convex effect. However, this effect becomes 0 and insignificant when controlling for country of origin in model (6). Additional controls for municipality in model (7) give a positive, concave yet insignificant effect of co-ethnic concentration on the log of hourly earnings. This could be due to the fact that certain municipalities dominated by certain sectors where wages tend to be lower experience higher concentration of co-ethnics, and higher concentration of co-ethnics in these municipalities in turn drives wages further down, so we cannot identify the temporality/directionality of this relationship.

### Issues  
1. There's not much variation in country of origin for high co-ethnic concentration levels (multicollinearity). Alternatively put, the most highly "ethnic" provinces are represented mainly by a few of the largest sending countries (Romania, Morocco, Ecuador). Thus, we don't know if we are controlling for country of origin or co-ethnic density, which are measured at different levels. 

2. When we control for country of origin in model (6) is when the effects of Spanish ability start to make sense, since Spanish ability and country of origin are correlated. Being a native speaker can be almost perfectly predicted from country of origin, so thus the negative effect of being a native speaker seen in models (3), (4), and (5) represents more accurately country of origin.

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.width=10,fig.height=7}
## Which countries are most represented?
countries <- table(this_pop$country) %>% as.data.frame(responseName = "Freq") %>% dplyr::rename(country = Var1) %>%
  arrange(desc(Freq))
these_top_countries <- countries %>% filter(Freq > 110)
this_pop <- this_pop %>% mutate(country_class = ifelse(country %in% these_top_countries$country, country, "Other"))
order_countries <- these_top_countries$country[these_top_countries$country %in% levels(factor(this_pop$country))] %>% as.vector()


ggplot(this_pop, aes(x=eth_dens_prov, y = log_hour_earn)) +
  geom_smooth(aes(x=eth_dens_prov, y = log_hour_earn), color = "black") +
  geom_point(aes(color=country_class)) + 
  geom_smooth(aes(color=country_class)) +
  scale_color_brewer(palette="Paired") + 
    labs(title = "Log of estimated hourly earnings on co-ethnic density \n Top-represented countries",
       x = "Co-ethnic density (province-level)",
       y = "Log of estimated hourly earnings") 
```


```{r echo = FALSE, warning = FALSE, message = FALSE, fig.width=10,fig.height=7}
ggplot(this_pop, aes(x=eth_dens_prov, fill=country_class)) +
  geom_histogram(bins = 52) + 
  scale_fill_brewer(palette="Paired") + 
    labs(title = "Counts of immigrants by co-ethnic density \n Top-represented countries",
       x = "Co-ethnic density (province-level)")
```


```{r echo = FALSE, warning = FALSE, message = FALSE, eval = F}
######### 2b. Effect of number of siblings living in same province on estimated hourly earnings
this_pop_2b <- this_pop %>% filter(!is.na(log_hour_earn))

model_2bi <- lm(log_hour_earn ~ HH_same_prov, data = this_pop_2b)

#model_2bi <- lm(log_hour_earn ~ HH_same_prov + HH_same_prov_sq, data = this_pop_2b)
#+ country_short + years + age + age_sq + sex + education + spanish + comunidad + occupation1 + work_permit

model_2bii <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish, data = this_pop_2b)

model_2biii <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit, data = this_pop_2b)

model_2biv <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex, data = this_pop_2b)

model_2bv <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short, data = this_pop_2b)

model_2bvi <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality, data = this_pop_2b)

model_2bvii <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1, data = this_pop_2b)

model_2bviii <- lm(log_hour_earn ~ HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1 + occupation1, data = this_pop_2b)

```

```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', eval = F}
stargazer(model_2bi, model_2bii, model_2biii, model_2biv, model_2bv, model_2bvi,
          type="latex",
          title = "Effect of siblings living in same province but not in same household on log of estimated hourly earnings",
          colnames = FALSE,
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Count of siblings in same province", "Years since arrival", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Work authorization?", "Age", "Age squared", "Sex - female"),
          keep = ,
          omit = c("country_short", "municipality"),
          add.lines=list(c('Country of origin controls', rep("No", 4), "Yes", "Yes"),
                         c('Municipality controls', rep("No", 5), "Yes")),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = TRUE,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "footnotesize",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")
```


```{r echo = FALSE, warning = FALSE, message = FALSE}
######### 3b. Effect of concentration of coethnics and siblings living in same province on estimated hourly earnings
this_pop_3b <- this_pop %>% filter(!is.na(log_hour_earn) & !is.na(eth_dens_prov))

model_3bi <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov, data = this_pop_3b)
#+ country_short + years + age + age_sq + sex + education + spanish + comunidad + occupation1 + work_permit

model_3bii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov+ years + education + spanish, data = this_pop_3b)

model_3biii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit, data = this_pop_3b)

model_3biv <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex, data = this_pop_3b)

model_3bv <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short, data = this_pop_3b)

model_3bvi <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality, data = this_pop_3b)

model_3bvii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1, data = this_pop_3b)

model_3bviii <- lm(log_hour_earn ~ eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + country_short + municipality + sector1 + occupation1, data = this_pop_3b)
```

```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', resize.height=4,resize.width=11}
stargazer(model_3bi, model_3bii, model_3biii, model_3biv, model_3bv, model_3bvi,
          type="latex",
          title = "Effect of concentration of coethnics and siblings living in same province on log of estimated hourly earnings",
          colnames = FALSE,
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Co-ethnic concentration", "Co-ethnic concentration squared", "Count of siblings in same province", "Years since arrival", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Work authorization?", "Age", "Age squared", "Sex - female"),
          keep = ,
          omit = c("country_short", "municipality"),
          add.lines=list(c('Country of origin controls', rep("No", 4), "Yes", "Yes"),
                         c('Municipality controls', rep("No", 5), "Yes")),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = TRUE,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "footnotesize",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")
```

## Table 2

### Observations
We observe a pretty consistent, slightly negative or zero-magntude and insignificant effect of the number of siblings one has living in the same province but not in the same household.  Individuals from certain countries of origin who earn less money per hour (negative effect) possess more "social capital" as measured by the number of siblings living in the same province, and when country of origin is controlled for we see this negative effect be removed when we move from model (4) to model (5). Certain municipalities with greater earnings per hour (most likely the urban ones, though I currently can't tell since I cannot link the municipality names to their codings) also have more siblings living in the same province as would make sense in the urban setting. When municipality is controlled for, we see this positive effect removed from the siblings variable as we move from model (5) to model (6).

### Issues
Number of siblings one has living in the same province is a very rudimentary and non-informative measure of social networks. There's not much variation and there is a lot of other factors affecting this measure that can also be related to wages. The idea of counting siblings *not living in the same household* was to try and address this potential endogeneity, as we could imagine that siblings in the household doesn't measure social networks but, say, more mouths to feed or greater sharing of financial resources within the household. Or perhaps less wages thus creates a greater need to live together and share resources. However, this does not avoid the fact that perhaps by doing particularly well economically, an individual spreads the word of his success to his family members or can then help pay the trip of a relative, prompting more to immigrate and thus an increase in number of siblings living in the same province. Perhaps one way to deal with this could be to look at “initial” social capital (how many family members or social support was accessed upon arrival) and restrict to those who didn’t have any pre-contract set up before arriving.

```{r echo = FALSE, warning = FALSE, message = FALSE}
######### 4b. Effect of concentration of coethnics in same province on estimated hourly earnings
this_pop_4b <- this_pop %>% filter(!is.na(log_hour_earn) & !is.na(participation_indicator) & spanish_nat == 0)
model_4bi <- lm(log_hour_earn ~ participation_indicator, data = this_pop_4b)
#+ country_short + years + age + age_sq + sex + education + spanish + comunidad + occupation1 + work_permit

model_4bii <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov, data = this_pop_4b)

model_4biii <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish, data = this_pop_4b)

model_4biv <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit, data = this_pop_4b)

model_4bv <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex, data = this_pop_4b)

model_4bvi <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + sex*participation_indicator, data = this_pop_4b)

model_4bvii <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + sex*participation_indicator + country_short, data = this_pop_4b)

model_4bviii <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + sex*participation_indicator+ country_short + municipality, data = this_pop_4b)


model_4bix <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + sex*participation_indicator + country_short + municipality + sector1, data = this_pop_4b)

model_4bx <- lm(log_hour_earn ~ participation_indicator + eth_dens_prov + eth_dens_prov_sq + HH_same_prov + years + education + spanish + work_permit + age + age_sq + sex + sex*participation_indicator + country_short + municipality + sector1 + occupation1, data = this_pop_4b)
```

```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', resize.height=4,resize.width=11}
stargazer(model_4bi, model_4bii, model_4biii, model_4biv, model_4bv, model_4bvi, model_4bvii, model_4bviii,
          type="latex",
          title = "Effect of civic participation on log of estimated hourly earnings (for those who do not possess Spanish nationality)",
          colnames = FALSE,
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Civic engagement indicator", "Co-ethnic concentration", "Co-ethnic concentration squared","Count of siblings in same province", "Years since arrival", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Work authorization?", "Age", "Age squared", "Sex - female", "Civic engagement:Sex - female"),
          keep = ,
          omit = c("country_short", "municipality"),
          add.lines=list(c('Country of origin controls', rep("No", 6), rep("Yes", 2)),
                         c('Municipality controls', rep("No", 7), rep("Yes", 1))),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = TRUE,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "tiny",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")
```

&nbsp;

## Table 3

### Observations
In general, we see a positive and significant effect of civic engagement on wages. The high magnitude effect (on the order of a ~10.9% average increase in wages for those who have participated in a local organization or voted in a municipal election versus those who haven't) decreases as we control for more covariates. For example, moving from model (2) to model (3) when control for education and years since arrival are added, we see that the ones who are highly educated and have been in Spain more years are more likely to have participated in a local organization or voted in municipal elections, hence the decrease in the positive effect of civic engagement. When an interaction term is added between civic engagement and gender, we see a significant and large additive effect of civic engagement for women. In model (6) of  Table 4 where controls for country of origin, municipality, sector and occupation of job in home country are added, we see that the average effect of civic engagement for women is an additional ~11.5% of estimated monthly earnings than for a similar man, while the effect of civic engagement for men is now negative (but insignificant). 

### Issues
With the civic engagement measure, we are not sure the *directionality* of the effect on wages. It could be that someone who is earning more is more likely to then go and participate in these organizations. Perhaps it costs money to join these organizations. It also requires time, which someone might not have much of if they have to work various jobs or are unable to pay other people to do housework/nannying for them (which is often work assigned to women), for example. It could also be a case of *simultaneity*. Perhaps a **structural equation modeling (SEM)** approach could be appropriate here. Finally, it should be investigated whether this civic engagement measure could be measuring other things that aren't related to social capital, per say. One example could be that some of these community organizations are actually helping immigrants validate their education certificates, and thus this assimilation of education obtained in the home country is what is causing the increase in wages.


```{r echo = FALSE, warning = FALSE, message = FALSE, results='asis', resize.height=4,resize.width=11}
stargazer(model_1bviii, model_1bix, model_3bvii, model_3bviii, model_4bix, model_4bx,
          object.names = F,
          column.labels = c("Table 1 cont.", "Table 1 cont.", "Table 2 cont.", "Table 2 cont.", "Table 3 cont.", "Table 3 cont."),
          type="latex",
          title = "Effect of concentration of coethnics and siblings living in same province and civic engagement on log of estimated hourly earnings \n Additional controls for sector and occupation in home country",
          colnames = FALSE,
          dep.var.labels = "Log of estimated hourly earnings",
          dep.var.labels.include = T,
          covariate.labels = c("Civic engagement", "Co-ethnic concentration", "Co-ethnic concentration squared", "Count of siblings in same province", "Years since arrival", "Education", "Spanish - suficiente", "Spanish - bien", "Spanish - nativo", "Work authorization?", "Age", "Age squared", "Sex - female", "Civic engagement:Sex - female"),
          keep = ,
          omit = c("country_short", "municipality", "sector1", "occupation1"),
          add.lines=list(c('Country of origin controls', rep("Yes", 6)),
                         c('Municipality controls', rep("Yes", 6)),
                         c('Sector controls', rep("Yes", 6)),
                         c('Occupation controls', rep(c("No", "Yes"), 3))),
          column.sep.width = "1pt",
          header=FALSE,
          no.space = TRUE,
          digits = 3,
          df = FALSE,
          single.row = T,
          font.size = "tiny",
          notes.align = "c",
          keep.stat = c("n", "rsq", "f", "ser"),
          table.placement = "H")

#"Sector - servicios empresariales", "Sector - servicio social", "Sector - administración pública", "Sector #- agricultura", "Sector - comercio", "Sector - construcción", "Sector - educacción", "Sector - hostelería", #"Sector - manufacturera", "Sector - extractiva", "Sector - financiera", "Sector - #extraterritoriales","Sector - servicios personales", "Sector - pesca", "Sector - energía", "Sector - #transporte",
```

## General comments

### Variable fixes
1. Need to work on the **co-ethnic concentration measure**. Right now, it is measured as the percentage of the total registered population in each of the 52 Spanish provinces that are from a specific country of origin in 2006. As of now, we have municipal-level data by country and sex from 2003 onwards, but there is an issue with matching the municipalities from the padrón data to the ENI data that needs to be resolved first. Additionally, at what point in time should we be measuring co-ethnic concentration to represent social capital? Should it be an average of the years someone lived in a specific province/municipality? What if someone just recently moved to the province/municipality at the end of 2006? Any effect on wages wouldn't be due to this co-ethnic concentration. *Perhaps co-ethnic concentration effect is better used to predict time it takes to find a job or hours worked*.  

2. **Level of Spanish**, which is hard to measure, might be better to capture Spanish ability through years lived in Spain and country of origin. Also, strong ability to speak Spanish could also partially measure social capital -- one who is able to communicate clearly with another and create mutual understanding builds trust.

3. Control for previous work experience: using sector and occupation of work in the country of origin are not the best way. Maybe it could be an indicator variable of whether someone has previous work experience in the same sector in which they're working in currently (whether **SECT2** matches with **SECT1** (sector of work in home country) or with **SECT3** (sector of work in first job in Spain).

4. Investigate potential measurement bias with the response variable, **log of estimate hourly earnings**. Perhaps there is seasonality issues with wage data and hours worked as data was collected from November 2006 to February 2007. This could specifically affect those working in, say, education, seasonal farm work, the tourism industry. Perhaps a control for **sector in which one currently works (SECT2)** could help.

### Future research
1. Read more to see how I might construct a better measure of social capital to address structural dynamics/directionality. Would it be possible to do something like Sabatini did in Italy with PCA and SEM?
    * Investigate the **TERESP** and **CONVA** variables to see whether civic engagement indicator is actually measuring whether someone convalidated their titles/certificates in Spain. 
2. Look into other response variables -- social networks and unemployment and labor mobility?
    * For those who have a lot of ethnic social support upon arriving (but no pre-arranged job), how likely were they to find a first job quickly or be in PARO?
    * For those who got their first job through contacts, how likely were they to upgrade their jobs
3. Investigate bridging vs. bonding social capital
    * Look at the within-household structure and determine if immigrants are living with mainly family members, other co-ethnics, or Spanish nationals (we have birth country, nationality(ies), age, sex, year and month of arrival to Spain ) Could potentially filter to childless immigrants and determine if they have some personal connection to a Spanish national (whether via partners, friends, housemates).


